### 2019-02-28 ms.Rmd
Start conversion to drake plan

---
title: Growth Dynamics of the Independent Gametophytes of *Pleurorosiopsis makinoi* (Polypodiaceae)
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load packages
library(conflicted)
library(tidyverse)
library(drake)
library(here)
library(cowplot)
library(ggplot2)
library(scales)
library(here)
library(patchwork)
library(lubridate)
library(gt)
library(ggridges)

# Resolve conflicts
col_factor <- conflict_prefer("col_factor", "readr")
discard    <- conflict_prefer("discard", "purrr")
expand     <- conflict_prefer("expand", "tidyr")
filter     <- conflict_prefer("filter", "dplyr")
gather     <- conflict_prefer("gather", "tidyr")
ggsave     <- conflict_prefer("ggsave", "ggplot2")
here       <- conflict_prefer("here", "here")
intersect  <- conflict_prefer("intersect", "dplyr")
lag        <- conflict_prefer("lag", "dplyr")
Position   <- conflict_prefer("Position", "ggplot2")
setdiff    <- conflict_prefer("setdiff", "dplyr")
union      <- conflict_prefer("union", "dplyr")

source(here("code/functions.R"))

#' Set order of microclimate variables for printing.
order_vars <- c("par_total", "rh_min", "temp_mean")

#' Select microclimate variables of interest for analysis.
selected_vars <- c("rh_min", "par_total", "temp_mean")

# Load objects from drake workflow into memory
loadd(combined_morph)
loadd(daily_microclimate)
loadd(t_test_results)
loadd(paired_microclimate)
loadd(nested_paired_microclimate)
```

```{r fig-3, fig.width=8, fig.height=7}
#' Set x-limits manually since the automatically set limits will vary
#' when we subset the data.
start_date <- combined_morph %>%
  pull(date) %>%
  min

end_date <- combined_morph %>%
  pull(date) %>%
  max

#' Make tibble of years for shading. This dataframe will be passed to `geom_rect()` when
#' plotting.
shading_dates <- 
  tibble(date = seq.Date(from = start_date, to = end_date, by = "day")) %>%
  mutate(year = year(date)) %>%
  # shade by alternate years
  filter(year %in% c(2009, 2011, 2013)) %>%
  group_by(year) %>%
  summarize(
    year_start = min(date),
    year_end = max(date)
  )

#' Make list of subplots. Build each one at a time, then add common
#' formatting. 
subplots <- list()

subplots[["cover"]] <- 
  combined_morph %>%
  ungroup %>%
  select(starts_with("q"), date) %>%
  filter(complete.cases(.)) %>%
  gather(plot, area, -date) %>%
  ggplot(aes(x = date, y = area, linetype = plot)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_linetype_manual(
    name="Quadrat",
    breaks=c("q_1", "q_2", "q_3", "q_4"),
    labels=c("1", "2", "3", "4"),
    values=c("solid", "dashed", "dotted", "dotdash")
  ) +
  labs(
    x = "",
    y = expression("Cover ("~cm^2~")"),
    subtitle = "a"
  ) +
  theme(legend.position = "bottom")

legend <- get_legend(subplots[["cover"]])

subplots[["cover"]] <- 
  subplots[["cover"]] +
  theme(legend.position = "none")

subplots[["gemmae_count"]] <- 
  combined_morph %>%
  select(count_mean, count_sd, date) %>%
  drop_na %>%
  ggplot(aes(x = date, y = count_mean)) %>%
  shade_years(shading_dates) +
  geom_errorbar(
    aes(ymin=count_mean-count_sd, 
        ymax=count_mean+count_sd), 
    width=.1,
    color = "dark grey") +
  geom_line() +
  geom_point(color = "black") +
  labs(
    x = "",
    y = expression("Gemmae count"),
    subtitle = "b"
  )

subplots[["gemmae_length"]] <- 
  combined_morph %>%
  select(length_mean, length_sd, date) %>%
  drop_na %>%
  ggplot(aes(x = date, y = length_mean)) %>%
  shade_years(shading_dates) +
  geom_errorbar(
    aes(ymin=length_mean-length_sd, 
        ymax=length_mean+length_sd), 
    width=.1,
    color = "dark grey") +
  geom_line() +
  geom_point(color = "black") +
  labs(
    x = "",
    y = expression(paste("Gemmae len. (", mu, "m)")),
    subtitle = "c"
  )

subplots <- subplots %>%
  map(~ . + 
        theme(
          axis.text.x = element_text(
            angle = 30, 
            hjust = 1, 
            vjust = 0.5, 
            margin=margin(-10,0,0,0)
          ),
          plot.margin = margin(0,10,0,10)
        ) +
        scale_x_date(
          date_labels = "%b %Y",
          date_breaks = "6 months",
          limits = c(start_date, end_date)
        )
  )

subplots[1:2] <- subplots[1:2] %>%
  map(~ . + theme(axis.text.x = element_blank()) 
  )

subplots[2:3] <- subplots[2:3] %>%
  map(~ . + theme(plot.margin = margin(-10,10,0,10)) 
  )

#' Combine subplots and write out
subplots[[1]] + subplots[[2]] + subplots[[3]] + 
  legend +
  plot_layout(ncol = 1, heights = c(1,1,1,0.2))

```

```{r fig-4, fig.width=8, fig.height=7}

#' Extract common start and end dates so x-axis are same across subplots.
start_date <- daily_microclimate %>% pull(date) %>% min
end_date <- daily_microclimate %>% pull(date) %>% max
all_days <- tibble( 
  date = seq(start_date, end_date, by = "day")
)

#' Make tibble of years for shading.
shading_dates <- 
  tibble(date = seq.Date(from = start_date, to = end_date, by = "day")) %>%
  mutate(year = year(date)) %>%
  # shade by odd years
  filter(year %in% c(2009, 2011, 2013)) %>%
  group_by(year) %>%
  summarize(
    year_start = min(date),
    year_end = max(date)
  )

#' PAR differs enough between sites that automatically set y-axis limits
#' would be different. Use the common min and max between them.
min_par <- daily_microclimate %>% pull(par_total) %>% min
max_par <- daily_microclimate %>% pull(par_total) %>% max

#' Insert NAs into microclimate for each site so that missing days
#' don't get connected by lines.
daily_microclimate_okutama <- filter(daily_microclimate, site == "okutama") %>%
  right_join(all_days)

daily_microclimate_takao <- filter(daily_microclimate, site == "uratakao") %>%
  right_join(all_days)

daily_microclimate_with_na <- bind_rows(
  daily_microclimate_okutama, daily_microclimate_takao)

#' Make list of subplots.
subplots <- list()

subplots[[1]] <- 
  daily_microclimate_okutama %>%
  ggplot(aes(x = date, y = par_total)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_y_continuous(limits = c(min_par, max_par)) +
  labs(y = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = "") ),
       x = "",
       title = "Okutama",
       subtitle = "a")

subplots[[2]] <- 
  daily_microclimate_takao %>%
  ggplot(aes(x = date, y = par_total)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_y_continuous(limits = c(min_par, max_par)) +
  labs(y = "",
       x = "",
       title = "Uratakao",
       subtitle = "b")

subplots[[3]] <- 
  daily_microclimate_okutama %>%
  ggplot(aes(x = date, y = rh_min)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(y = "Min. Rel. Humidity (%)",
       x = "",
       subtitle = "c")

subplots[[4]] <- 
  daily_microclimate_takao %>%
  ggplot(aes(x = date, y = rh_min)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(y = "",
       x = "",
       subtitle = "d")

subplots[[5]] <- 
  daily_microclimate_okutama %>%
  ggplot(aes(x = date, y = temp_mean)) %>%
  shade_years(shading_dates) +
  geom_line() +
  labs(y = "Mean Temp. (°C)",
       x = "",
       subtitle = "e")

subplots[[6]] <- 
  daily_microclimate_takao %>%
  ggplot(aes(x = date, y = temp_mean)) %>%
  shade_years(shading_dates) +
  geom_line() +
  labs(y = "",
       x = "",
       subtitle = "f")

#' Apply common formatting.
subplots <- subplots %>%
  map(~ . + 
        theme(
          axis.text.x = element_text(
            angle = 30, 
            hjust = 1, 
            vjust = 0.5, 
            margin=margin(-10,0,0,0)
          )
        ) +
        scale_x_date(
          date_labels = "%b %Y",
          date_breaks = "6 months",
          limits = c(start_date, end_date)
        )
  )

#' Remove x-axis lables from upper plots.
subplots[1:4] <- subplots[1:4] %>%
  map(~ . + theme(axis.text.x = element_blank()) 
  )

#' Close gaps on top margins for lower plots.
subplots[c(3,5)] <- subplots[c(3,5)] %>%
  map(~ . + theme(plot.margin = margin(-10,10,0,10)) 
  )

#' Close gaps on left side for RHS plots.
subplots[c(2,4,6)] <- subplots[c(2,4,6)] %>%
  map(~ . + theme(plot.margin = margin(-10,10,0,-40)) 
  )

#' Combine plots and write out.
subplots[[1]] + subplots[[2]] + subplots[[3]] + 
  subplots[[4]] + subplots[[5]] + subplots[[6]] + 
  plot_layout(ncol = 2)

```

```{r table-1}
#' Format Table 1 and output to RTF using `gt` package.
t_test_results %>%
  arrange(factor(var, levels = order_vars)) %>%
  mutate(season = stringr::str_to_title(season)) %>%
  mutate(var = case_when(
    var == "rh_min" ~    "Min. Rel. Humidity",
    var == "par_total" ~ "DLI",
    var == "temp_mean" ~ "Mean Temp."
  )) %>%
  group_by(season) %>%
  gt() %>%
  fmt_number(
    columns = vars(estimate, statistic, conf.low, conf.high, 
                   okutama_mean, okutama_sd, 
                   uratakao_mean, uratakao_sd)
  ) %>%
  cols_move(
    columns = vars(okutama_mean, okutama_sd),
    after = vars(conf.high)
  ) %>%
  tab_spanner(
    label = "Okutama",
    columns = vars(okutama_mean, okutama_sd)
  ) %>%
  tab_spanner(
    label = "Uratakao",
    columns = vars(uratakao_mean, uratakao_sd)
  ) %>%
  cols_label(
    var = "",
    statistic = "t",
    pval = "p",
    parameter = "df",
    conf.low = "Lower",
    conf.high = "Upper",
    okutama_mean = "Mean",
    okutama_sd = "SD",
    uratakao_mean = "Mean",
    uratakao_sd = "SD"
  ) %>%
  tab_style(
    style = cells_styles(
      bkgd_color = "white"),
    locations = cells_data(
      columns = everything()
    )
  ) %>%
  tab_header(
    title = "Table 1"
  )

```

```{r fig-5, fig.width=7, fig.height=8}
#' Extract p-values for plotting, and put into data frame
#' with xvals specifying season and yvals specifying height 
#' to plot asterisks.
p_vals <-
  t_test_results %>%
  mutate(pval = case_when(
    pval < 0.0001 ~ "***",
    pval < 0.001 ~ "**",
    pval < 0.05 ~ "*",
    TRUE ~ ""
  )) %>%
  select(season, var, pval) %>%
  spread(var, pval) %>%
  rename(
    par_total_pval = par_total,
    rh_min_pval = rh_min,
    temp_mean_pval = temp_mean
  )

# Simple function to calculate maximum y-position for plotting asterisks.
# Should be just above the mean + error bar height (1 sd).
maxy <- function(x) {mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)}

pval_plot_data <- 
  paired_microclimate %>%
  select(-date, -month) %>%
  group_by(season, site) %>%
  summarize_all(
    maxy
  ) %>%
  group_by(season) %>%
  summarize_all(
    max
  ) %>%
  select(-site) %>%
  left_join(p_vals)

#' Format data for plotting: need columns to be mean and sd, by site and season.
paired_means <- 
  paired_microclimate %>%
  select(-date) %>%
  group_by(site, season) %>%
  summarize_all(
    list(~mean(., na.rm = TRUE),
         ~sd(., na.rm = TRUE))
  ) %>%
  mutate(
    season = factor(season, levels = c("winter", "spring", "summer", "fall"))
  )

#' Make plot including t-test on three microclimatic variables.
#' 
#' The final plot includes multiple subplots. Each is a ridge plot showing 
#' the distribution of one of the variables between the two sites, with the
#' results of a paired t-test showing if they are different or not.
#' 
#' Need to manually set x ranges to keep x-axis scale the same across plots
x_ranges <- tibble(var = selected_vars) %>%
  mutate(
    min = map_dbl(var, ~ pull(paired_microclimate, .) %>% min ),
    max = map_dbl(var, ~ pull(paired_microclimate, .) %>% max )
  )

# light blue / light green colorblind-friendly from colorbrewer
# paired_colors <- c("#a6cee3", "#b2df8a")
# OR dark blue / dark green colorblind-friendly from colorbrewer
# paired_colors <- c("#1f78b4", "#33a02c")
# OR dark grey / light grey for bw
paired_colors <- c("grey55", "grey90")

subplots <- list()

#' Make ridge plots by mapping across nested datasets.
plot_results <-
  nested_paired_microclimate %>%
  right_join(x_ranges) %>%
  mutate(min = min - 0.05 * min,
         max = max + 0.05 * max,
         plot = pmap(list(var, data, x_min = min, x_max = max), 
                     plot_ridges, 
                     y_var = "site"),
         pval = map2(var, data, run_t_test),
         pval = case_when(
           pval < 0.0001 ~ "***",
           pval < 0.001 ~ "**",
           pval < 0.05 ~ "*",
           TRUE ~ ""
         ),
         plot = map2(
           plot, pval, 
           ~ .x + annotate("text", Inf, Inf, 
                           label = .y, hjust = 1, vjust = 1))

         # If needed, add annotation for season to make sure subplots
         # are positioned correctly

         # ,plot = map2(
         #   plot, season,
         #   ~ .x + annotate("text", -Inf, Inf,
         #                   label = .y, hjust = -1, vjust = 1)),
         # season = as_factor(season),
         # season = fct_relevel(season, "winter", "spring", "summer", "fall")

  ) %>% 
  arrange(season, var)

subplots <- plot_results$plot

#' Manually some subplots to remove redundant labels, etc.
subplots[1:9] <- map(subplots[1:9], 
                     ~ . + theme(axis.title.x = element_blank(),
                                 axis.text.x = element_blank() ))

subplots[1:12] <- map(subplots[1:12], 
                      ~ . + scale_fill_manual(
                        values = paired_colors,
                        name="Site",
                        breaks=c("okutama", "uratakao"),
                        labels=c("Okutama", "Uratakao"))
)

subplots[[10]] <- subplots[[10]] +
  labs(x = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = "")) )

subplots[[11]] <- subplots[[11]] +
  labs(x = expression("Min. Rel. Humidity (%)"))

subplots[[12]] <- subplots[[12]] +
  labs(x = expression("Mean Temp. (°C)"))

#' Add titles.
subplots <- map(subplots, ~ . + labs(title = "") )

subplots[[2]] <- subplots[[2]] +
  labs(title = "Winter\n")

subplots[[5]] <- subplots[[5]] +
  labs(title = "Spring\n")

subplots[[8]] <- subplots[[8]] +
  labs(title = "Summer\n")

subplots[[11]] <- subplots[[11]] +
  labs(title = "Fall\n")

#' Extract legend
legend <- paired_microclimate %>%
  ggplot(aes(x = temp_mean, y = site, fill = site)) +
  geom_density_ridges() +
  theme(legend.position = "bottom") +
  scale_fill_manual(
    values = paired_colors,
    name="Site",
    breaks=c("okutama", "uratakao"),
    labels=c("Okutama", "Uratakao"))

legend <- get_legend(legend)

#' Combine and write out.
subplots[[1]] +  subplots[[2]] +  subplots[[3]] + 
  subplots[[4]] +  subplots[[5]] +  subplots[[6]] + 
  subplots[[7]] +  subplots[[8]] +  subplots[[9]] + 
  subplots[[10]] + subplots[[11]] + subplots[[12]] +
  plot_spacer() + legend + plot_spacer() + plot_layout(ncol = 3, heights = c(1, 1, 1, 1, 0.2))
```