### 2019-03-01 ms.Rmd
Finish running all analyses and output to ms as drake plan

---
title: Growth Dynamics of the Independent Gametophytes of *Pleurorosiopsis makinoi* (Polypodiaceae)
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Set order of microclimate variables for printing.
order_vars <- c("par_total", "rh_min", "temp_mean")

# Load objects from drake workflow into memory
loadd(selected_vars)
loadd(combined_morph)
loadd(daily_microclimate)
loadd(t_test_results)
loadd(paired_microclimate)
loadd(nested_paired_microclimate)
loadd(combined_monthly_morph)
loadd(linear_model_results)
loadd(gameto_size)
```

```{r fig-3, fig.width=8, fig.height=7}
#' Set x-limits manually since the automatically set limits will vary
#' when we subset the data.
start_date <- combined_morph %>%
  pull(date) %>%
  min

end_date <- combined_morph %>%
  pull(date) %>%
  max

#' Make tibble of years for shading. This dataframe will be passed to `geom_rect()` when
#' plotting.
shading_dates <- 
  tibble(date = seq.Date(from = start_date, to = end_date, by = "day")) %>%
  mutate(year = year(date)) %>%
  # shade by alternate years
  filter(year %in% c(2009, 2011, 2013)) %>%
  group_by(year) %>%
  summarize(
    year_start = min(date),
    year_end = max(date)
  )

#' Make list of subplots. Build each one at a time, then add common
#' formatting. 
subplots <- list()

subplots[["cover"]] <- 
  combined_morph %>%
  ungroup %>%
  select(starts_with("q"), date) %>%
  filter(complete.cases(.)) %>%
  gather(plot, area, -date) %>%
  ggplot(aes(x = date, y = area, linetype = plot)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_linetype_manual(
    name="Quadrat",
    breaks=c("q_1", "q_2", "q_3", "q_4"),
    labels=c("1", "2", "3", "4"),
    values=c("solid", "dashed", "dotted", "dotdash")
  ) +
  labs(
    x = "",
    y = expression("Cover ("~cm^2~")"),
    subtitle = "a"
  ) +
  theme(legend.position = "bottom")

legend <- get_legend(subplots[["cover"]])

subplots[["cover"]] <- 
  subplots[["cover"]] +
  theme(legend.position = "none")

subplots[["gemmae_count"]] <- 
  combined_morph %>%
  select(count_mean, count_sd, date) %>%
  drop_na %>%
  ggplot(aes(x = date, y = count_mean)) %>%
  shade_years(shading_dates) +
  geom_errorbar(
    aes(ymin=count_mean-count_sd, 
        ymax=count_mean+count_sd), 
    width=.1,
    color = "dark grey") +
  geom_line() +
  geom_point(color = "black") +
  labs(
    x = "",
    y = expression("Gemmae count"),
    subtitle = "b"
  )

subplots[["gemmae_length"]] <- 
  combined_morph %>%
  select(length_mean, length_sd, date) %>%
  drop_na %>%
  ggplot(aes(x = date, y = length_mean)) %>%
  shade_years(shading_dates) +
  geom_errorbar(
    aes(ymin=length_mean-length_sd, 
        ymax=length_mean+length_sd), 
    width=.1,
    color = "dark grey") +
  geom_line() +
  geom_point(color = "black") +
  labs(
    x = "",
    y = expression(paste("Gemmae len. (", mu, "m)")),
    subtitle = "c"
  )

subplots <- subplots %>%
  map(~ . + 
        theme(
          axis.text.x = element_text(
            angle = 30, 
            hjust = 1, 
            vjust = 0.5, 
            margin=margin(-10,0,0,0)
          ),
          plot.margin = margin(0,10,0,10)
        ) +
        scale_x_date(
          date_labels = "%b %Y",
          date_breaks = "6 months",
          limits = c(start_date, end_date)
        )
  )

subplots[1:2] <- subplots[1:2] %>%
  map(~ . + theme(axis.text.x = element_blank()) 
  )

subplots[2:3] <- subplots[2:3] %>%
  map(~ . + theme(plot.margin = margin(-10,10,0,10)) 
  )

#' Combine subplots and write out
subplots[[1]] + subplots[[2]] + subplots[[3]] + 
  legend +
  plot_layout(ncol = 1, heights = c(1,1,1,0.2))

```

```{r fig-4, fig.width=8, fig.height=7}

#' Extract common start and end dates so x-axis are same across subplots.
start_date <- daily_microclimate %>% pull(date) %>% min
end_date <- daily_microclimate %>% pull(date) %>% max
all_days <- tibble( 
  date = seq(start_date, end_date, by = "day")
)

#' Make tibble of years for shading.
shading_dates <- 
  tibble(date = seq.Date(from = start_date, to = end_date, by = "day")) %>%
  mutate(year = year(date)) %>%
  # shade by odd years
  filter(year %in% c(2009, 2011, 2013)) %>%
  group_by(year) %>%
  summarize(
    year_start = min(date),
    year_end = max(date)
  )

#' PAR differs enough between sites that automatically set y-axis limits
#' would be different. Use the common min and max between them.
min_par <- daily_microclimate %>% pull(par_total) %>% min
max_par <- daily_microclimate %>% pull(par_total) %>% max

#' Insert NAs into microclimate for each site so that missing days
#' don't get connected by lines.
daily_microclimate_okutama <- filter(daily_microclimate, site == "okutama") %>%
  right_join(all_days)

daily_microclimate_takao <- filter(daily_microclimate, site == "uratakao") %>%
  right_join(all_days)

daily_microclimate_with_na <- bind_rows(
  daily_microclimate_okutama, daily_microclimate_takao)

#' Make list of subplots.
subplots <- list()

subplots[[1]] <- 
  daily_microclimate_okutama %>%
  ggplot(aes(x = date, y = par_total)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_y_continuous(limits = c(min_par, max_par)) +
  labs(y = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = "") ),
       x = "",
       title = "Okutama",
       subtitle = "a")

subplots[[2]] <- 
  daily_microclimate_takao %>%
  ggplot(aes(x = date, y = par_total)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_y_continuous(limits = c(min_par, max_par)) +
  labs(y = "",
       x = "",
       title = "Uratakao",
       subtitle = "b")

subplots[[3]] <- 
  daily_microclimate_okutama %>%
  ggplot(aes(x = date, y = rh_min)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(y = "Min. Rel. Humidity (%)",
       x = "",
       subtitle = "c")

subplots[[4]] <- 
  daily_microclimate_takao %>%
  ggplot(aes(x = date, y = rh_min)) %>%
  shade_years(shading_dates) +
  geom_line() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(y = "",
       x = "",
       subtitle = "d")

subplots[[5]] <- 
  daily_microclimate_okutama %>%
  ggplot(aes(x = date, y = temp_mean)) %>%
  shade_years(shading_dates) +
  geom_line() +
  labs(y = "Mean Temp. (°C)",
       x = "",
       subtitle = "e")

subplots[[6]] <- 
  daily_microclimate_takao %>%
  ggplot(aes(x = date, y = temp_mean)) %>%
  shade_years(shading_dates) +
  geom_line() +
  labs(y = "",
       x = "",
       subtitle = "f")

#' Apply common formatting.
subplots <- subplots %>%
  map(~ . + 
        theme(
          axis.text.x = element_text(
            angle = 30, 
            hjust = 1, 
            vjust = 0.5, 
            margin=margin(-10,0,0,0)
          )
        ) +
        scale_x_date(
          date_labels = "%b %Y",
          date_breaks = "6 months",
          limits = c(start_date, end_date)
        )
  )

#' Remove x-axis lables from upper plots.
subplots[1:4] <- subplots[1:4] %>%
  map(~ . + theme(axis.text.x = element_blank()) 
  )

#' Close gaps on top margins for lower plots.
subplots[c(3,5)] <- subplots[c(3,5)] %>%
  map(~ . + theme(plot.margin = margin(-10,10,0,10)) 
  )

#' Close gaps on left side for RHS plots.
subplots[c(2,4,6)] <- subplots[c(2,4,6)] %>%
  map(~ . + theme(plot.margin = margin(-10,10,0,-40)) 
  )

#' Combine plots and write out.
subplots[[1]] + subplots[[2]] + subplots[[3]] + 
  subplots[[4]] + subplots[[5]] + subplots[[6]] + 
  plot_layout(ncol = 2)

```

```{r table-1}
#' Format Table 1 and output to RTF using `gt` package.
t_test_results %>%
  arrange(factor(var, levels = order_vars)) %>%
  mutate(season = stringr::str_to_title(season)) %>%
  mutate(var = case_when(
    var == "rh_min" ~    "Min. Rel. Humidity",
    var == "par_total" ~ "DLI",
    var == "temp_mean" ~ "Mean Temp."
  )) %>%
  group_by(season) %>%
  gt() %>%
  fmt_number(
    columns = vars(estimate, statistic, conf.low, conf.high, 
                   okutama_mean, okutama_sd, 
                   uratakao_mean, uratakao_sd)
  ) %>%
  cols_move(
    columns = vars(okutama_mean, okutama_sd),
    after = vars(conf.high)
  ) %>%
  tab_spanner(
    label = "Okutama",
    columns = vars(okutama_mean, okutama_sd)
  ) %>%
  tab_spanner(
    label = "Uratakao",
    columns = vars(uratakao_mean, uratakao_sd)
  ) %>%
  cols_label(
    var = "",
    statistic = "t",
    pval = "p",
    parameter = "df",
    conf.low = "Lower",
    conf.high = "Upper",
    okutama_mean = "Mean",
    okutama_sd = "SD",
    uratakao_mean = "Mean",
    uratakao_sd = "SD"
  ) %>%
  tab_style(
    style = cells_styles(
      bkgd_color = "white"),
    locations = cells_data(
      columns = everything()
    )
  ) %>%
  tab_header(
    title = "Table 1"
  )

```

```{r fig-5, evaluate = FALSE, fig.width=7, fig.height=8}
# #' Extract p-values for plotting, and put into data frame
# #' with xvals specifying season and yvals specifying height 
# #' to plot asterisks.
# p_vals <-
#   t_test_results %>%
#   mutate(pval = case_when(
#     pval < 0.0001 ~ "***",
#     pval < 0.001 ~ "**",
#     pval < 0.05 ~ "*",
#     TRUE ~ ""
#   )) %>%
#   select(season, var, pval) %>%
#   spread(var, pval) %>%
#   rename(
#     par_total_pval = par_total,
#     rh_min_pval = rh_min,
#     temp_mean_pval = temp_mean
#   )
# 
# # Simple function to calculate maximum y-position for plotting asterisks.
# # Should be just above the mean + error bar height (1 sd).
# maxy <- function(x) {mean(x, na.rm = TRUE) + sd(x, na.rm = TRUE)}
# 
# pval_plot_data <- 
#   paired_microclimate %>%
#   select(-date, -month) %>%
#   group_by(season, site) %>%
#   summarize_all(
#     maxy
#   ) %>%
#   group_by(season) %>%
#   summarize_all(
#     max
#   ) %>%
#   select(-site) %>%
#   left_join(p_vals)
# 
# #' Format data for plotting: need columns to be mean and sd, by site and season.
# paired_means <- 
#   paired_microclimate %>%
#   select(-date) %>%
#   group_by(site, season) %>%
#   summarize_all(
#     list(~mean(., na.rm = TRUE),
#          ~sd(., na.rm = TRUE))
#   ) %>%
#   mutate(
#     season = factor(season, levels = c("winter", "spring", "summer", "fall"))
#   )
# 
# #' Make plot including t-test on three microclimatic variables.
# #' 
# #' The final plot includes multiple subplots. Each is a ridge plot showing 
# #' the distribution of one of the variables between the two sites, with the
# #' results of a paired t-test showing if they are different or not.
# #' 
# #' Need to manually set x ranges to keep x-axis scale the same across plots
# x_ranges <- tibble(var = selected_vars) %>%
#   mutate(
#     min = map_dbl(var, ~ pull(paired_microclimate, .) %>% min ),
#     max = map_dbl(var, ~ pull(paired_microclimate, .) %>% max )
#   )
# 
# # light blue / light green colorblind-friendly from colorbrewer
# # paired_colors <- c("#a6cee3", "#b2df8a")
# # OR dark blue / dark green colorblind-friendly from colorbrewer
# # paired_colors <- c("#1f78b4", "#33a02c")
# # OR dark grey / light grey for bw
# paired_colors <- c("grey55", "grey90")
# 
# subplots <- list()
# 
# #' Make ridge plots by mapping across nested datasets.
# plot_results <-
#   nested_paired_microclimate %>%
#   right_join(x_ranges) %>%
#   mutate(min = min - 0.05 * min,
#          max = max + 0.05 * max,
#          plot = pmap(list(var, data, x_min = min, x_max = max), 
#                      plot_ridges, 
#                      y_var = "site"),
#          pval = map2(var, data, run_t_test),
#          pval = case_when(
#            pval < 0.0001 ~ "***",
#            pval < 0.001 ~ "**",
#            pval < 0.05 ~ "*",
#            TRUE ~ ""
#          ),
#          plot = map2(
#            plot, pval, 
#            ~ .x + annotate("text", Inf, Inf, 
#                            label = .y, hjust = 1, vjust = 1))
#          
#          # If needed, add annotation for season to make sure subplots
#          # are positioned correctly
#          
#          # ,plot = map2(
#          #   plot, season,
#          #   ~ .x + annotate("text", -Inf, Inf,
#          #                   label = .y, hjust = -1, vjust = 1)),
#          # season = as_factor(season),
#          # season = fct_relevel(season, "winter", "spring", "summer", "fall")
#          
#   ) %>% 
#   arrange(season, var)
# 
# subplots <- plot_results$plot
# 
# #' Manually some subplots to remove redundant labels, etc.
# subplots[1:9] <- map(subplots[1:9], 
#                      ~ . + theme(axis.title.x = element_blank(),
#                                  axis.text.x = element_blank() ))
# 
# 
# subplots[1:12] <- map(subplots[1:12], 
#                       ~ . + scale_fill_manual(
#                         values = paired_colors,
#                         name="Site",
#                         breaks=c("okutama", "uratakao"),
#                         labels=c("Okutama", "Uratakao"))
# )
# 
# subplots[[10]] <- subplots[[10]] +
#   labs(x = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = "")) )
# 
# subplots[[11]] <- subplots[[11]] +
#   labs(x = expression("Min. Rel. Humidity (%)"))
# 
# subplots[[12]] <- subplots[[12]] +
#   labs(x = expression("Mean Temp. (°C)"))
# 
# #' Add titles.
# subplots <- map(subplots, ~ . + labs(title = "") )
# 
# subplots[[2]] <- subplots[[2]] +
#   labs(title = "Winter\n")
# 
# subplots[[5]] <- subplots[[5]] +
#   labs(title = "Spring\n")
# 
# subplots[[8]] <- subplots[[8]] +
#   labs(title = "Summer\n")
# 
# subplots[[11]] <- subplots[[11]] +
#   labs(title = "Fall\n")
# 
# #' Extract legend
# legend <- paired_microclimate %>%
#   ggplot(aes(x = temp_mean, y = site, fill = site)) +
#   geom_density_ridges() +
#   theme(legend.position = "bottom") +
#   scale_fill_manual(
#     values = paired_colors,
#     name="Site",
#     breaks=c("okutama", "uratakao"),
#     labels=c("Okutama", "Uratakao"))
# 
# legend <- get_legend(legend)
# 
# #' Combine and write out.
# subplots[[1]] +  subplots[[2]] +  subplots[[3]] + 
#   subplots[[4]] +  subplots[[5]] +  subplots[[6]] + 
#   subplots[[7]] +  subplots[[8]] +  subplots[[9]] + 
#   subplots[[10]] + subplots[[11]] + subplots[[12]] +
#   plot_spacer() + legend + plot_spacer() + plot_layout(ncol = 3, heights = c(1, 1, 1, 1, 0.2))
```

```{r table-2}
cols_numeric <- c("r.squared", "adj.r.squared", "sigma", "statistic", "logLik", "AIC", "BIC")

linear_model_results %>%
  select(-fits) %>%
  arrange(factor(x_var, levels = order_vars)) %>%
  mutate(
    x_var = case_when(
      x_var == "rh_min" ~    "Min. Rel. Humidity",
      x_var == "par_total" ~ "DLI",
      x_var == "temp_mean" ~ "Mean Temp."
    ),
    y_var = case_when(
      y_var == "total_cover" ~ "Total cover",
      y_var == "count_mean" ~ "Gemmae count",
      y_var == "length_mean" ~ "Gemmae length"
    )
  ) %>%
  group_by(x_var) %>%
  gt() %>%
  cols_move_to_start(vars(y_var)) %>%
  cols_move(
    columns = vars(df.residual),
    after = vars(df)
  ) %>% 
  fmt_number(
    columns = colnames(.) %in% cols_numeric
  ) %>%
  fmt_number(
    columns = vars(p.value),
    decimals = 3
  ) %>%
  fmt_number(
    columns = vars(deviance),
    decimals = 0
  ) %>%
  cols_label(
    y_var = "",
    r.squared = "r-squared",
    adj.r.squared = "adj. r-squared",
    statistic = "t",
    p.value = "p",
    df.residual = "df residual"
  ) %>%
  tab_style(
    style = cells_styles(
      bkgd_color = "white"),
    locations = cells_data(
      columns = everything()
    )
  ) %>%
  tab_header(
    title = "Table 2"
  )

```

```{r fig-6, fig.width=7, fig.height=8}
#' Loop through all combination of dep and indep variables and their models,
#' and make plots for each.
var_grid <- make_var_grid (c("rh_min", "par_total", "temp_mean"))

subplots <- 
  pmap(var_grid, 
       quick_plot_with_stats, 
       plot_data = combined_monthly_morph, 
       model_data = linear_model_results)

#' Add axis titles
#' 
#' First, remove all default titles, then add formatted titles.
subplots <- map(subplots, ~ . + labs(x = "", y = ""))
subplots[[1]] <- subplots[[1]] + labs(y = expression("Total cover ("~cm^2~")"))
subplots[[4]] <- subplots[[4]] + labs(y = "Gemmae count")
subplots[[7]] <- subplots[[7]] + 
  labs(
    y = expression(paste("Gemmae len. (", mu, "m)", sep = "")),
    x = expression(paste("DLI (", mol~m^-2~day^-1, ")", sep = ""))
  )
subplots[[8]] <- subplots[[8]] + labs(x = expression("Min. Rel. Humidity (%)"))
subplots[[9]] <- subplots[[9]] + labs(x = expression("Mean Temp. (°C)"))

#' Set amount of gap to close on left and top
left_close <- -15
top_close <- -10

#' Close gaps on left only: 2,3
subplots[2:3] <- subplots[2:3] %>%
  map(~ . + theme(plot.margin = margin(0,0,0,left_close)) 
  )

#' Close gaps on left side and top: 5,6,8,9
subplots[c(5,6,8,9)] <- subplots[c(5,6,8,9)] %>%
  map(~ . + theme(plot.margin = margin(top_close,0,0,left_close)) 
  )

#' Close gaps on top only: 4,7
subplots[c(4,7)] <- subplots[c(4,7)] %>%
  map(~ . + theme(plot.margin = margin(top_close,0,0,0)) 
  )

#' Add legend for middle bottom plot, combine, and write out.
subplots[[8]] <- subplots[[8]] + 
  theme(legend.position="bottom") + 
  labs(color = "Month", shape = "Year")

subplots[[1]] + subplots[[2]] + subplots[[3]] + 
  subplots[[4]] + subplots[[5]] + subplots[[6]] + 
  subplots[[7]] + subplots[[8]] + subplots[[9]] + 
  plot_layout(ncol = 3)

#' Black and white version if needed.
# subplots <- 
#   map(subplots,
#       ~ . + scale_color_grey() 
#   )
# 
# subplots[[1]] + subplots[[2]] + subplots[[3]] + 
#   subplots[[4]] + subplots[[5]] + subplots[[6]] + 
#   subplots[[7]] + subplots[[8]] + subplots[[9]] + 
#   plot_layout(ncol = 3)
```

```{r fig-7, fig.width=7, fig.height=4}
#' Note there is one outlier with length 46 mm, 
#' then the rest all less than 19 mm. We will remove this 
#' before plotting and make a note in the figure caption.
gameto_size_no_outlier <-
  gameto_size %>%
  # Change sex to uppercase for nicer plotting
  mutate(sex = stringr::str_to_title(sex)) %>%
  # Remove one outlier
  filter(length < 40)

#' Make jitter plots overlayed with box plots. Make two
#' subplots: one for width and one for length, then combine.

# Use colorbrewer palette: 3 diverging, color-blind friendly colors
cols <- c("#1b9e77","#d95f02","#7570b3")

# Try to make sure dots stay in the same place, but doesn't work
set.seed(1000)

# Color version

subplots <- list()

subplots[["length"]] <- 
  ggplot(gameto_size_no_outlier, aes(y = length, x = sex, color = sex)) +
  labs(
    subtitle = "a",
    y = "Length (mm)",
    x = "",
    color = "")

subplots[["width"]] <-
  ggplot(gameto_size_no_outlier, aes(y = width, x = sex, color = sex)) +
  labs(
    subtitle = "b",
    y = "Width (mm)",
    x = "")

subplots <- 
  map(subplots,
      ~ . + 
        geom_jitter(size = 3) +
        geom_boxplot(
          outlier.shape = NA,
          fill = NA,
          color = "dark grey",
          alpha = 0.7
        ) +
        scale_color_manual(values = cols) +
        theme(legend.position = "none")
  )

subplots[[1]] + subplots[[2]] + plot_layout(ncol = 2)

# Black and white version
# subplots <- 
#   map(subplots,
#       ~ . + 
#         scale_color_grey() +
#         theme(legend.position = "none")
#   )

```